---
title: "Functions, subsetting, indexing extras"
author: "Bryan Shalloway"
date: "5/31/2019"
output: html_document
---

### Changing values with indexes

```{r}
library(tidyverse)
```


In this section I use the word 'indexes' to refer to any base R method for specifying position ^[I am using the word loosely to mean both either the situation when you specify postions by a series of `TRUE` / `FALSE` values or a series of numeric indexes E.g. x[c(TRUE, TRUE, FALSE, FALSE, TRUE)] or x[c(1, 2, 5)] , 'index' obviously sounds more like the later, but I mean it generally to cover just 'base R method of specifying positoin'.].

My solution to building the new function in 19.2.1.2 replaces values over re-writing a new vector, namely by specifying indexes and then forcing those to be a new value.  
For example, say we have a vector `c(1:20)` and we want to make all even values equal to 0.  below is how you could do that by simply re-writing this to a new vector
```{r}
x <- c(1:20)

x_0_even <- ifelse((x %% 2) == 0, 0, x)

x_0_even
```

Alternatively, you could do this by simply overwriting the values in a specified index with a value (like I did in question 19.2.1.2)
```{r}
x <- c(1:20)

x[(x %% 2) == 0] <- 0

x
```

Both the indexing and the ifelse method give the same output. I have a slight preference for the `ifelse` method as I think it is a little easier to read. Also, it doesn't force you to overwrite your data -- to save against this second problem I will often save a copy before applying the indexed approach (though if it's in a function don't need to worry about this as changes will default to occur within function not global environment). e.g.

```{r}
x <- c(1:20)
x_0_even <- x

x_0_even[(x %% 2) == 0] <- 0

x_0_even
```

If you're curious about speed of each, you can see the index method tends to be faster on this dataset.  

*Build functions (necessary for measuring speed): *
```{r}
method_ifelse <- function(vector = c(1:20)){
  x <- vector
  
  x_0_even <- ifelse((x %% 2) == 0, 0, x)
  
  x_0_even
}

method_index <- function(vector = c(1:20)){
  x <- vector
  x[(x %% 2) == 0] <- 0
  
  x
}
```

*Measure time it takes to run: *
```{r}
microbenchmark::microbenchmark(ifelse = method_ifelse(1:1000),
                               index = method_index(1:1000),
                               times = 500)

```

The index methods tends to be faster.

#### Applying indexing to dfs

I have a high preference for using tidyverse style approaches when applying transformations to dataframes, though there are instances when it's easier to use indexing methods. A common example of this is say we want to replace all of the `NA` values across multiple columns in a dataframe.

```{r}
df <- tibble(x = c(NA, 3, 4),
       y = c(4, NA, NA))

```

The code below uses an indexing method to replace all `NA` values in the df with 0
```{r}
df_cleanNA <- df
df_cleanNA[is.na(df)] <- 0

df_cleanNA
```

Below is the dplyr approach (requires knowledge of `mutate_all` and `funs`)
```{r}
mutate_all(df, funs(ifelse(is.na(.), 0, .)))
```

With a purrr function you could have done:
```{r}
purrr::map_df(df, ~ifelse(is.na(.x), 0, .x))

```

If you're curious of speed of each, below is microbenchmark test

```{r}
df_na0_index <- function(df){
  df[is.na(df)] <- 0
  df
}

df_na0_dplyr <- function(df){
  mutate_all(df, funs(ifelse(is.na(.), 0, .)))
}

df_na0_purrr <- function(df){
  purrr::map_df(df, ~ifelse(is.na(.x), 0, .x))
}
```

*Measure time it takes to run: *  

First on tiny dataset:
```{r}
microbenchmark::microbenchmark(index = df_na0_index(flights),
                               dplyr = df_na0_index(flights),
                               purrr = df_na0_purrr(flights),
                               times = 10)
```


Then on larger dataset:
```{r}
microbenchmark::microbenchmark(index = df_na0_index(flights),
                               dplyr = df_na0_index(flights),
                               purrr = df_na0_purrr(flights),
                               times = 10)
```

You should see that the index and dplyr method are pretty consistent on time, whereas the purrr method will be slower. For this example and some problems like this then indexing may be the best option (generally though I lean more towards using dplyr where possible).  

I also wonder if there may be a better alternative to `ifelse` in this situation.

### Better than `ifelse()`?

I end-up using `ifelse` a lot for basic transformations, I'm curious if there is a more efficient alternative for use with dplyr style... 

* I learned about dplyr's functions `dplyr::recode` which is similar to `forcats::fct_recode` and can be used for replacing multiple character values. 
* `dplyr::case_when` can be used for more complex criteria
* `dplyr::if_else` is similar to `base::ifelse` but more strict on types

### Dplyr and functions

As was mentioned, dplyr uses non-standard evaluation. This means that when referring to column names form within a function, dplyr will require some slightly different syntax... I typically find ways around this by deploying tricks with the `*_` verbs or taking advantage of the `vars` functions and `*_at` or by messing with the names, though am interested to hear other methods...

Say you want to write a function that takes in a dataframe and a list of column names and you want to return a sum of these into a new column with the name corresponding with the name of the spliced together columns. What would be the most elegant way of doing this with tidyverse style?
