---
title: "explanation of the `with` function"
date: "`r paste('Last updated: ', format(Sys.time(), '%Y-%m-%d'))`"
author: "Bryan Shalloway"
output: 
  github_document:
    toc: true
    toc_depth: 3
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE)

```


```{r, warning = FALSE}
library(tidyverse)
library(nycflights13)
```

The `with` function takes a named list or dataframe as its first argument. The second argument can be an an expression you want evaluated. `with` then allows you to type column names to insert the associated vector into that location. The ability to evaluate expressions by just referencing the column names is called "non-standard evaluation" very similar to how dplyr facilitates this (also using "non-standard evaluation").

For example, say I want to output the column `Sepal.Width` from the `iris` dataset as a vector.

```{r}
iris <- as_tibble(iris)
```


```{r}
# base R methods
iris$Sepal.Width
iris[["Sepal.Width"]]

# using the with function
with(iris, Sepal.Width)

# the above are the same, just different methods

# note that the below dplyr function gives you something different
# the output is a one-column dataframe as opposed to an atomic vector
select(iris, Sepal.Width)
```

Say I want to take the `sum` of `Sepal.Width` and ouptut it as a vector. `with` makes it easier to keep this in a pipe format.
```{r}
iris %>% 
  with(sum(Sepal.Width))
```

You can also reference more than one column vector in an expression. E.g. lets look at using it with the `table` function (`table` essentially works like a pivot-table with default count option).
```{r}
flights %>% 
  with(table(carrier, origin))

```

Or maybe I want to use `table` and then pass the outputs into `cor` (to check the correlation of flights by carrier between the three airports in NYC).

```{r}
 flights %>% 
  with(table(carrier, origin)) %>% 
  cor()
```

* LGA and JFK seem to have at least some correlation of flights and carrier

Essentially, I use `with` when I want to do operations in a "non-standard evaluation" format similar to that used in dplyr while also facilitating the use of the pipe. I'm mostly working in dataframes so I don't use it a ton, though sometimes `with` can be very useful.
