*Make sure the following packages are installed:*  

```{r setup10, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

library(ggplot2)
library(dplyr)
library(nycflights13)
library(readr)
```

# ch. 10: Tibbles

```{r, eval = FALSE}
vignette("tibble")
```

* `tibble`: produces a dataframe w/ some other helpful qualities that have advantages over `data.frame`
* `as_tibble`: convert to a tibble
* `tribble`: transposed tibble - set-up for data entry into a tibble in code
* `print`: can use print to set how the tibble will print
```{r}
nycflights13::flights %>% 
  print(n = 2, width = Inf)
```

    + Also can convert with `as.data.frame` or use `options`, see 10.5.6 below
* `enframe`: let's you encode name and value, see 10.5.5 below
* `class`: for checking the class of the object
    + Though is not fully accurate, in that the actual object class of vectors is "base", not double, etc., so kind of lies...

## 10.5

*1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).*  
  
Could look at printing, e.g. only prints first 15 rows and enough variables where you can see them all, or by checking explicitly the `class` function^[or could check a few other things such as if list-cols are supported]

*2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?*  
  
Dataframes can't do list-cols. Never changes type of input e.g. from strings to factors, never changes names of variables, never creates row names. Also, you can do list-cols with tibbles.  
  

*3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?*  
  
```{r, eval = FALSE}
var <- "var_name"

# Will extract the column as an atomic vector
df[[var]]
```

*4. Practice referring to non-syntactic names in the following data frame by:*  
  
```{r}
df <- tibble(`1` = 1:10, `2` = 11:20)

```

*a. Extracting the variable called 1.*  
```{r}
df %>% 
  select(1)
```
  

*b. Plotting a scatterplot of 1 vs 2.*  
```{r}
df %>% 
  ggplot(aes(x = `1`, y = `2`))+
  geom_point()
```
 
*c. Creating a new column called 3 which is 2 divided by 1.*  
  
```{r}
df %>% 
  mutate(`3` = `1` / `2`) 
```

*d. Renaming the columns to one, two and three.*  
```{r}
df %>% 
  mutate(`3` = `1` / `2`) %>% 
  rename(one = `1`,
         two = `2`,
         three = `3`)

```
  

*5. What does tibble::enframe() do? When might you use it?*  

Let's you encode "name" and "value"
```{r}
tibble::enframe(1:3)
tibble::enframe(c(a = 5, b = 8))
tibble::enframe(c(a = 5:8, b = 7:10))
tibble::enframe(c(a = 5:8, b = 7:10, d = 9:12))
```

*6. What option controls how many additional column names are printed at the footer of a tibble?*  

* argument `tibble.width`

```{r, eval = FALSE}
options(tibble.print_max = n, tibble.print_min = m)
options(tibble.width = Inf)
options(dplyr.print_min = Inf) #to always show all rows
```
