*Make sure the following packages are installed:*  

```{r setup15, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(nycflights13)
library(lubridate)
library(forcats)
```

# ch. 15: Factors

* `factor` make variable a factor based on `levels` provided
* `fct_rev` reverses order of factors
* `fct_infreq` orders levels in increasing frequency
* `fct_relevel` lets you move levels to front of order
* `fct_inorder` orders existing factor by order values show-up in in data
* `fct_reorder` orders input factors by other specified variables value (median by default), 3 inputs: `f`: factor to modify, `x`: input var to order by, `fun`: function to use on x, also have `desc` option
* `fct_reorder2` orders input factor by max of other specified variable (good for making legends align as expected)
* `fct_recode` lets you change value of each level
* `fct_collapse` is variant of `fct_recode` that allows you to provide multiple old levels as a vector
* `fct_lump` allows you to lump together small groups, use `n` to specify number of groups to end with

Create factors by order they come-in:

Avoiding dropping levels with `drop = FALSE`
```{r}
gss_cat %>% 
  ggplot(aes(race))+
  geom_bar()+
  scale_x_discrete(drop = FALSE)

# Is there a similar way to do this with count?
```

## 15.4: Modifying factor order

Example with `fct_recode`

```{r}
gss_cat %>%
  mutate(partyid = fct_recode(partyid,
    "Republican, strong"    = "Strong republican",
    "Republican, weak"      = "Not str republican",
    "Independent, near rep" = "Ind,near rep",
    "Independent, near dem" = "Ind,near dem",
    "Democrat, weak"        = "Not str democrat",
    "Democrat, strong"      = "Strong democrat"
  )) %>%
  count(partyid)
```


### 15.3.1 

1.  Explore the distribution of `rincome` (reported income). What makes the
    default bar chart hard to understand? How could you improve the plot?
    
    ```{r}
    gss_cat %>% 
      ggplot(aes(x = rincome)) +
      geom_bar()

    ```
    
    
    * Default bar chart has categories across the x-asix, I flipped these to be across the y-axis 
    * Also, have highest values at the bottom rather than at the top and have different version of NA showing-up at both top and bottom, all should be on one side
    * In `bar_prep`, I used reg expressions to extract the numeric values, arrange by that, and then set factor levels according to the new order
        + Solution is probably unnecessarily complicated...^[Also had issue with not rendering for book.]

    ```{r, error = TRUE}
    bar_prep <- gss_cat %>% 
      tidyr::extract(col = rincome, into =c("dollars1", "dollars2"), "([0-9]+)[^0-9]*([0-9]*)", remove = FALSE) %>% 
      mutate_at(c("dollars1", "dollars2"), ~ifelse(is.na(.) | . == "", 0, as.numeric(.))) %>% 
      arrange(dollars1, dollars2) %>% 
      mutate(rincome = fct_inorder(rincome))
    
    bar_prep %>% 
      ggplot(aes(x = rincome)) +
      geom_bar() +
      scale_x_discrete(drop = FALSE) +
      coord_flip()
    ```


1.  What is the most common `relig` in this survey? What's the most
    common `partyid`?

    ```{r}
    gss_cat %>%
      count(relig, sort = TRUE)
      
    gss_cat %>%
      count(partyid, sort = TRUE)
      
    ```

    * `relig` most common -- Protestant, 10846,
    * `partyid` most common -- Independent, 4119

1.  Which `relig` does `denom` (denomination) apply to? How can you find
    out with a table? How can you find out with a visualisation?
    
    *With visualization:*
    ```{r}
    gss_cat %>% 
      ggplot(aes(x=relig, fill=denom))+
      geom_bar()+
      coord_flip()
    ```
    
    * Notice which have the widest variety of colours -- are protestant, and Christian slightly
    
    *With table:*
    ```{r}
    gss_cat %>% 
      count(relig, denom) %>% 
      count(relig, sort = TRUE)
    ```
    
	
## 	15.4: Modifying factor order

### 15.4.1
	
1.  There are some suspiciously high numbers in `tvhours`. Is the mean a good
    summary?
    
    ```{r}
    gss_cat %>%
      mutate(tvhours_fct = factor(tvhours)) %>% 
      ggplot(aes(x = tvhours_fct)) +
      geom_bar()
    
    ```
   
   * Distribution is reasonably skewed with some values showing-up as 24 hours which seems impossible, in addition to this we have a lot of na values, this may skew results
   * Given high number of missing values, `tvhours` may also just not be reliable, do NA's associate with other variables? -- Perhaps could try and impute these NAs

1.  For each factor in `gss_cat` identify whether the order of the levels is
    arbitrary or principled.
    
    ```{r}
    gss_cat %>% 
      purrr::keep(is.factor) %>% 
      purrr::map(levels)
    
    ```
    
    * `rincome` is principaled, rest are arbitrary


1.  Why did moving "Not applicable" to the front of the levels move it to the
    bottom of the plot?
    
    * Becuase is moving this factor to be first in order
	
## 	15.5: Modifying factor levels

### 15.5.1

1.  How have the proportions of people identifying as Democrat, Republican, and
    Independent changed over time?

    *As a line plot: * 
    ```{r}
    gss_cat %>%
      mutate(partyid = fct_collapse(
        partyid,
        other = c("No answer", "Don't know", "Other party"),
        rep = c("Strong republican", "Not str republican"),
        ind = c("Ind,near rep", "Independent", "Ind,near dem"),
        dem = c("Not str democrat", "Strong democrat")
      )) %>%
      count(year, partyid) %>%
      group_by(year) %>%
      mutate(prop = n / sum(n)) %>%
      ungroup() %>%
      ggplot(aes(
        x = year,
        y = prop,
        colour = fct_reorder2(partyid, year, prop)
      )) +
      geom_line() +
      labs(colour = "partyid")
    ```
    
    *As a bar plot: * 
    
    ```{r}
    gss_cat %>%
      mutate(partyid = fct_collapse(
        partyid,
        other = c("No answer", "Don't know", "Other party"),
        rep = c("Strong republican", "Not str republican"),
        ind = c("Ind,near rep", "Independent", "Ind,near dem"),
        dem = c("Not str democrat", "Strong democrat")
      )) %>%
      count(year, partyid) %>%
      group_by(year) %>%
      mutate(prop = n / sum(n)) %>%
      ungroup() %>%
      ggplot(aes(
        x = year,
        y = prop,
        fill = fct_reorder2(partyid, year, prop)
      )) +
      geom_col() +
      labs(colour = "partyid")
    
    ```

    * Suggests proportion of republicans has gone down with independents and other going up.

1.  How could you collapse `rincome` into a small set of categories?

    ```{r}
    other = c("No answer", "Don't know", "Refused", "Not applicable")
    high = c("$25000 or more", "$20000 - 24999", "$15000 - 19999", "$10000 - 14999")
    med = c("$8000 to 9999", "$7000 to 7999", "$6000 to 6999", "$5000 to 5999")
    low = c("$4000 to 4999", "$3000 to 3999", "$1000 to 2999", "Lt $1000")
    
    mutate(gss_cat,
           rincome = fct_collapse(
           rincome,
           other = other,
           high = high,
           med = med,
           low = low
           )) %>%
           count(rincome)
    
    ```



# Appendix 

## Viewing all levels

A few ways to get an initial look at the levels or counts across a dataset

```{r}
gss_cat %>% 
  purrr::map(unique)

gss_cat %>% 
  purrr::map(table)

gss_cat %>% 
  purrr::map(table) %>% 
  purrr::map(plot)

```

## Percentage NA each level

```{r}
gss_cat %>% 
  purrr::map(~(sum(is.na(.x)) / length(.x)))
```

## Print all levels of tibble

```{r}
gss_cat %>% 
  count(tvhours) %>% 
  print(n = Inf)
```

