
```{r}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE)
```

*Make sure the following packages are installed:*  
```{r setup18}
library(ggplot2)
library(dplyr)
library(forcats)
library(tidyr)
library(lubridate)
library(stringr)
```


# ch. 18: Pipes (notes only)

* `pryr::object_size` gives the memory occupied by all of its arguments (note that built-in object.size does not allow measuring multiple objects so can't see shared space). This function is actually shown in chapter 18: Pipes
* Some functions do not work naturally with the pipe.
    + If you want to use `assign` with the pipe, you must be explicit about the environment

```{r}
env <- environment()
"x" %>% assign(100, envir = env)
x
```

* `try`, `tryCatch`, `suppressMessages`, and `suppressWarnings` from base R all also do not work well

Other pipes = 'T pipe', `%T>%` that returns left-hand side rather than right.  Will let the plot output, but then continues.  Notice that this doesn't work for ggplot as ggplot does output something

```{r}
library(magrittr)
rnorm(100) %>%
  matrix(ncol = 2) %T>%
  plot() %>% 
  str()

iris %>% 
  select(Sepal.Length, Sepal.Width) %T>%  
  plot() %>% 
  select(Sepal.Length) %>% 
  head(10)

```

* `%$%` allows you to blow out the names of the arguments, I personally prefer using the `with` function for this instead as I find it to be a little more readable...
  + The two examples below are equivalent
```{r}
mtcars %$%
  cor(disp, mpg)

mtcars %>%
  with(cor(disp, mpg))

```

