---
title: "Chapters 1 to 3"
date: "`r paste('Last updated: ', format(Sys.time(), '%Y-%m-%d'))`"
author: "Bryan Shalloway"
output: 
  github_document:
    toc: true
    toc_depth: 3
---
 

```{r, include = FALSE, eval = FALSE}
rmarkdown::render("ch1to3.Rmd", output_format = "all")
```


*Make sure the following packages are installed:*  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

library(ggplot2)
library(maps)
library(mapproj)
library(dplyr)
```

# Chapter 3

## 3.2: First steps
### 3.2.4
**1. Run ggplot(data = mpg). What do you see?**

```{r}
ggplot(data = mpg)
```

Just a blank grey space.

**2. How many rows are in mpg? How many columns?**

```{r}
ncol(mtcars)
nrow(mtcars)
```


**3. What does the drv variable describe? Read the help for ?mpg to find out.**  

Front wheel, rear wheel or 4 wheel drive.  

  
**4. Make a scatterplot of hwy vs cyl.**

```{r}
ggplot(mpg)+
  geom_point(aes(x = hwy, y = cyl))

```

Inverse relationship.  
  
**5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?**  
*(key question)*
```{r, 3.2.4.5}
ggplot(mpg)+
  geom_point(aes(x = class, y = drv))

```

The points stack-up on top of one another so you don't get a sense of how many are on each point.  

*Any ideas for what methods could you use to improve the view of this data?*

## 3.3: Aesthetic mappings
### 3.3.1.
**1. What’s gone wrong with this code? Why are the points not blue?**
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))

```
  
The `color` field is in the `aes` function so it is expecting a character or factor variable. By inputting "blue" here, ggplot reads this as a character field with the value "blue" that it then supplies it's default color schemes to (1st: salmon, 2nd: teal)

**2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?**  

```{r}
mpg
```

The data is in tibble form already so just printing it shows the type, but could also use the `glimpse` and `str` functions.


**3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?**
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy, color = cyl, size = displ, shape = fl))

```
  
`color`: For continuous applies a gradient, for categorical it applies distinct colors based on the number of categories.  
`size`: For continuous, applies in order, for categorical will apply in an order that may be arbitrary if there is not an order provided.  
`shape`: Will not allow you to input a continuous variable.  

**4. What happens if you map the same variable to multiple aesthetics?**  
Will map onto both fields. Can be redundant in some cases, in others it can be valuable for clarity.
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x = cty, y = hwy, color = fl, shape = fl))
```


**5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)**
```{r}
?geom_point

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point(shape = 21, colour = "black", fill = "white", size = 5, stroke = 5)

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point(shape = 21, colour = "black", fill = "white", size = 5, stroke = 3)

```

For shapes that have a border (like 21), you can colour the inside and outside separately. Use the stroke aesthetic to modify the width of the border.   

**6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?**
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy, colour = displ < 5)) +
  geom_point()
```

The field becomes a logical operator in this case.

## 3.5: Facets

### 3.5.1.

**1. What happens if you facet on a continuous variable?**

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy))+
  geom_point()+
  facet_wrap(~cyl)
```

It will facet along all of the possible values.  

**2. What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean?**  
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

**How do they relate to this plot?**
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```

They represent the locations where there is no point on the above graph (could be made more clear by giving consistent order to axes).

**3. What plots does the following code make? What does . do?**

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```


Can use to specify if to facet by rows or columns.  


**4. Take the first faceted plot in this section:**
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

**What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?**  
  
Faceting prevents overlapping points in the data. A disadvantage is that you have to move your eye to look at different graphs. Some groups you don't have much data on as well so those don't present much information. If there is more data, you may be more comfortable using facetting as each group should have points that you can view.


**5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn't facet_grid() have nrow and ncol argument?**  
  
`nrow` and `ncol` specify the number of columns or rows to facet by, `facet_grid` does not have this option because the splits are defined by the number of unique values in each variable. Other important options are `scales` which let you define if the scales are able to change between each plot.


**6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?**  
  
I'm not sure why exactly, if I compare these, it's not completely unclear.
```{r}
#more unique levels on columns
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy)) + 
  facet_grid(year ~ class)

#more unique levels on rows
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy)) + 
  facet_grid(class ~ year)
```

My guess though would be that it's because our computer screens are generally wider than they are tall. Hence there will be more space for viewing a higher number of attributes going across columns than by rows.  

## 3.6: Geometric Objects

### 3.6.1

**1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?**  
`geom_line`  
`geom_boxplot`  
`geom_histogram`  
`geom_area`  

*Example of `geom_area`*  
Notice that `geom_area` is just a special case of `geom_ribbon`
```{r}
huron <- data.frame(year = 1875:1972, level = as.vector(LakeHuron) - 575)
h <- ggplot(huron, aes(year))

h + geom_ribbon(aes(ymin = 0, ymax = level))
h + geom_area(aes(y = level))

# Add aesthetic mappings
h +
  geom_ribbon(aes(ymin = level - 1, ymax = level + 1), fill = "grey70") +
  geom_line(aes(y = level))

h +
  geom_area(aes(y = level), fill = "grey70") +
  geom_line(aes(y = level))

```



**2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.**  
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```


**3. What does show.legend = FALSE do? What happens if you remove it? Why do you think I used it earlier in the chapter?**  
```{r}
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```

It get's rid of the legend that would be assogiated with this geom. You removed it previously to keep it consistent with your other graphs that did not include them to specify the `drv`.  

**4. What does the `se` argument to `geom_smooth()` do?**  
`se` here stands for standard error, so if we specify it as `FALSE` we are saying we do not want to show the standard errors for the plot.  


**5. Will these two graphs look different? Why/why not?**  
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

No, because local mappings for each geom are the same as the global mappings in the other.  

**6. Recreate the R code necessary to generate the following graphs.**  

```{r}
ggplot(mpg, aes(displ, hwy))+
  geom_point() +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(mpg, aes(displ, hwy, group = drv))+
  geom_point() +
  geom_smooth(se = FALSE)

```

```{r}
ggplot(mpg, aes(displ, hwy, colour = drv))+
  geom_point() +
  geom_smooth(se = FALSE)

```

```{r}
ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(colour = drv)) +
  geom_smooth(se = FALSE)

```


```{r}
ggplot(mpg, aes(displ, hwy))+
  geom_point(aes(color = drv)) +
  geom_smooth(aes(linetype = drv), se = FALSE)

```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(colour = "white", size = 4) +
  geom_point(aes(colour = drv))

```

## 3.7: statistical transformations

### 3.7.1.

**1. What is the default `geom` associated with `stat_summary()`? How could you rewrite the previous plot to use that geom function instead of the stat function?**  

The default is `geom_pointrange`,  the point being the mean, and the lines being the standard error on the y value (i.e. the deviation of the mean of the value).
```{r}
ggplot(mpg) +
  stat_summary(aes(cyl, cty))

```

*Rewritten with geom^[See [3.7.1.1 extension] for notes on how to relate this to dplyr code.]:*

```{r}
ggplot(mpg)+
  geom_pointrange(aes(x = cyl, y = cty), stat = "summary")

```

The specific example though is actually not the default:

```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

*Rewritten with geom:*
```{r}
ggplot(data = diamonds)+
  geom_pointrange(aes(x = cut, y = depth), 
                  stat = "summary", 
                  fun.ymin = "min",
                  fun.ymax = "max", 
                  fun.y = "median")

```

**2. What does geom_col() do? How is it different to geom_bar()?**  
`geom_col` has `"identity" as the default `stat`, so it expects to receive a variable that already has the value aggregated^[I often use this over `geom_bar` and do the aggregation with dplyr rather than ggplot2]


**3.Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?**  

```{r}
?ggplot2
```


**4.What variables does stat_smooth() compute? What parameters control its behaviour?**  
See here: http://ggplot2.tidyverse.org/reference/#section-layer-stats for a helpful resource.
Also, someone who aggregated some online: http://sape.inf.usi.ch/quick-reference/ggplot2/geom ^[Though it's missing some very common ones like `geom_col` and `geom_bar`.]


**5. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?**  
(key question)
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```


```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))

```

This is a solution, but still seems off as prop becomes out of a value greater than 1
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop.., group = color))
```

For this second graph though, I would think you would want something more like the following:
```{r}

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color), position = "fill")

```

Which could be generated by this code as well
```{r, eval = FALSE}
diamonds %>% 
  count(cut, color) %>% 
  group_by(cut) %>% 
  mutate(prop = n / sum(n)) %>% 
  ggplot(aes(x = cut, y = prop, fill = color))+
  geom_col()

```

## 3.8: Position Adjjustment

### 3.8.1.

**1.What is the problem with this plot? How could you improve it?**
(key question)
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```

The points overlap, could use `geom_jjitter` instead

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter()
```

**2. What parameters to geom_jitter() control the amount of jittering?**  
`height` and `width`

**3. Compare and contrast geom_jitter() with geom_count().**  
(key question)
Take the above chart and instead use `geom_count`
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count()

```

Can also use `geom_count` with `color`, and can use "jitter" in `position` arg.
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy, colour = drv)) + 
  geom_count()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy, colour = drv)) + 
  geom_count(position = "jitter")


ggplot(data = mpg, mapping = aes(x = cty, y = hwy, colour = drv)) + 
  geom_jitter(size = 3, alpha = 0.3)

```

One problem with `geom_count` is that the shapes can still block-out other shapes at that same point of different colors. You can flip the orderof the stacking order of the colors with `position` = "dodge". Still this seems limited.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy, colour = drv)) + 
  geom_count(position = "dodge")
```


**4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.**  
`dodge`, but seems like `identity` is the same

```{r}
ggplot(data=mpg, mapping=aes(x=class, y=hwy))+
  geom_boxplot()

```

## 3.9: Coordinate systems

### 3.9.1.

**1.Turn a stacked bar chart into a pie chart using `coord_polar()`.**  
These are more illustrative than anything, here is a note from the documetantion:  
*NOTE: Use these plots with caution - polar coordinates has major perceptual problems. The main point of these examples is to demonstrate how these common plots can be described in the grammar.  Use with EXTREME caution.*
```{r}
ggplot(mpg, aes(x = 1, fill = class))+
  geom_bar(position = "fill") +
  coord_polar(theta = "y") + 
  scale_x_continuous(labels = NULL)
```

If I want to make multiple levels:
```{r}
ggplot(mpg, aes(x = as.factor(cyl), fill = class))+
  geom_bar(position = "fill") +
  coord_polar(theta = "y")
```


**2. What does labs() do? Read the documentation.**  
Used for giving labels.
```{r}
?labs
```


**3. What’s the difference between `coord_quickmap()` and `coord_map()`?**  
The first is an approximation, useful for smaller regions to be proected. For this example, do not see substantial differences.
```{r}
nz <- map_data("nz")

ggplot(nz,aes(long,lat,group=group))+
  geom_polygon(fill="red",colour="black")+
  coord_quickmap()

ggplot(nz,aes(long,lat,group=group))+
  geom_polygon(fill="red",colour="black")+
  coord_map()


```


**4. What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?**  
`geom_abline()` adds a line with a given intercept and slope (either given by `aes` or by `intercept` and `slope` args)  
`coord_fixed` ensures that the ratios between the x and y axis stay at a specified relationship (default = 1). This is important for easily seeing the magnitude of the relationship between variables.  

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

# Appendix
## 3.7.1.1 extension
```{r}
ggplot(mpg, aes(x = cyl, y = cty, group = cyl))+
  geom_pointrange(stat = "summary")

```

This seems to be the same as what you would get by doing the following with dplyr:
```{r}
mpg %>% 
  group_by(cyl) %>% 
  dplyr::summarise(mean = mean(cty),
            sd = (sum((cty - mean(cty))^2) / (n() - 1))^0.5,
            n = n(),
            se = sd / n^0.5,
            lower = mean - se,
            upper = mean + se) %>% 
  ggplot(aes(x = cyl, y = mean, group = cyl))+
  geom_pointrange(aes(ymin = lower, ymax = upper))

```

Other geoms you could have set stat_summary to:  

`crossbar`:
```{r}
ggplot(mpg) +
  stat_summary(aes(cyl, cty), geom = "crossbar")

```

`errorbar`: 
```{r}
ggplot(mpg) +
  stat_summary(aes(cyl, cty), geom = "errorbar")

```

`linerange`:
```{r}
ggplot(mpg) +
  stat_summary(aes(cyl, cty), geom = "linerange")

```


## 3.8: Position adustments

Some "dodge"" examples
```{r}

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")

diamonds %>% 
  count(cut, color) %>% 
  ggplot(aes(x = cut, y = n, fill = color))+
  geom_col(position = "dodge")

```

Looking of `geom_jitter` and only changing width.
```{r}
ggplot(data = mpg, mapping = aes(x = drv, y = hwy))+
  geom_jitter(height = 0, width = .2)

```

## 3.9: Coordinate systems  
`coord_flip` is helpful, especially for quickly tackling issues with axis labels
`coord_quickmap` is important to remember if plotting spatial data.
`coord_polar` is important to remember if plotting spatial coordinates.
`map_data` for extracting data on maps of locations

## add in table of contents and other details...