---
title: 'Chapter 19: Functions'
date: "`r paste('Last updated: ', format(Sys.time(), '%Y-%m-%d'))`"
author: "Bryan Shalloway"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

*Make sure the following packages are installed:*  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(nycflights13)
library(babynames)
library(nasaweather)
library(lubridate)
```

# ch. 19: Functions

## 19.3: Functions are for humans and computers

### 19.3.1

1.  Read the source code for each of the following three functions, puzzle out
    what they do, and then brainstorm better names.
    
    ```{r}
    f1 <- function(string, prefix) {
      substr(string, 1, nchar(prefix)) == prefix
    }
    f2 <- function(x) {
      if (length(x) <= 1) return(NULL)
      x[-length(x)]
    }
    f3 <- function(x, y) {
      rep(y, length.out = length(x))
    }
    ```
    
1.  Take a function that you've written recently and spend 5 minutes 
    brainstorming a better name for it and its arguments.

1.  Compare and contrast `rnorm()` and `MASS::mvrnorm()`. How could you make
    them more consistent? 
    
1.  Make a case for why `norm_r()`, `norm_d()` etc would be better than
    `rnorm()`, `dnorm()`. Make a case for the opposite.
    
## 19.4: Conditional execution

### 19.4.4

1.  What's the difference between `if` and `ifelse()`? Carefully read the help
    and construct three examples that illustrate the key differences.

1.  Write a greeting function that says "good morning", "good afternoon",
    or "good evening", depending on the time of day. (Hint: use a time
    argument that defaults to `lubridate::now()`. That will make it 
    easier to test your function.)

1.  Implement a `fizzbuzz` function. It takes a single number as input. If
    the number is divisible by three, it returns "fizz". If it's divisible by
    five it returns "buzz". If it's divisible by three and five, it returns
    "fizzbuzz". Otherwise, it returns the number. Make sure you first write 
    working code before you create the function.
    
1.  How could you use `cut()` to simplify this set of nested if-else statements?

    ```{r, eval = FALSE}
    if (temp <= 0) {
      "freezing"
    } else if (temp <= 10) {
      "cold"
    } else if (temp <= 20) {
      "cool"
    } else if (temp <= 30) {
      "warm"
    } else {
      "hot"
    }
    ```
    
    How would you change the call to `cut()` if I'd used `<` instead of `<=`?
    What is the other chief advantage of `cut()` for this problem? (Hint:
    what happens if you have many values in `temp`?)

1.  What happens if you use `switch()` with numeric values?

1.  What does this `switch()` call do? What happens if `x` is "e"?

    ```{r, eval = FALSE}
    switch(x, 
      a = ,
      b = "ab",
      c = ,
      d = "cd"
    )
    ```
    
    Experiment, then carefully read the documentation. 
    
## 19.5: Function arguments

### 19.5.5

1.  What does `commas(letters, collapse = "-")` do? Why?

1.  It'd be nice if you could supply multiple characters to the `pad` argument, 
    e.g. `rule("Title", pad = "-+")`. Why doesn't this currently work? How 
    could you fix it?
    
1.  What does the `trim` argument to `mean()` do? When might you use it?

1.  The default value for the `method` argument to `cor()` is 
    `c("pearson", "kendall", "spearman")`. What does that mean? What 
    value is used by default?