---
title: "chapter 21: Iteration"
date: "`r paste('Last updated: ', format(Sys.time(), '%Y-%m-%d'))`"
author: "Bryan Shalloway"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

*Make sure the following packages are installed:*  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(nycflights13)
library(babynames)
library(nasaweather)
library(lubridate)
```

# ch. 21: Iteration

## 21.2: For loops

### 21.2.1

1.  Write for loops to:

    1. Compute the mean of every column in `mtcars`.
    1. Determine the type of each column in `nycflights13::flights`.
    1. Compute the number of unique values in each column of `iris`.
    1. Generate 10 random normals for each of $\mu = -10$, $0$, $10$, and $100$.
    
    Think about the output, sequence, and body __before__ you start writing
    the loop.

1.  Eliminate the for loop in each of the following examples by taking 
    advantage of an existing function that works with vectors:
    
    ```{r, eval = FALSE}
    out <- ""
    for (x in letters) {
      out <- stringr::str_c(out, x)
    }
    
    x <- sample(100)
    sd <- 0
    for (i in seq_along(x)) {
      sd <- sd + (x[i] - mean(x)) ^ 2
    }
    sd <- sqrt(sd / (length(x) - 1))
    
    x <- runif(100)
    out <- vector("numeric", length(x))
    out[1] <- x[1]
    for (i in 2:length(x)) {
      out[i] <- out[i - 1] + x[i]
    }
    ```

1.  Combine your function writing and for loop skills:

    1. Write a for loop that `prints()` the lyrics to the children's song 
       "Alice the camel".

    1. Convert the nursery rhyme "ten in the bed" to a function. Generalise 
       it to any number of people in any sleeping structure.

    1. Convert the song "99 bottles of beer on the wall" to a function.
       Generalise to any number of any vessel containing any liquid on 
       any surface.

1.  It's common to see for loops that don't preallocate the output and instead
    increase the length of a vector at each step:
    
    ```{r, eval = FALSE}
    output <- vector("integer", 0)
    for (i in seq_along(x)) {
      output <- c(output, lengths(x[[i]]))
    }
    output
    ```
    
    How does this affect performance? Design and execute an experiment.
    
## 21.3 For loop variations

### 21.3.5

1.  Imagine you have a directory full of CSV files that you want to read in.
    You have their paths in a vector, 
    `files <- dir("data/", pattern = "\\.csv$", full.names = TRUE)`, and now
    want to read each one with `read_csv()`. Write the for loop that will 
    load them into a single data frame. 

1.  What happens if you use `for (nm in names(x))` and `x` has no names?
    What if only some of the elements are named? What if the names are
    not unique?

1.  Write a function that prints the mean of each numeric column in a data 
    frame, along with its name. For example, `show_mean(iris)` would print:
    
    ```{r, eval = FALSE}
    show_mean(iris)
    #> Sepal.Length: 5.84
    #> Sepal.Width:  3.06
    #> Petal.Length: 3.76
    #> Petal.Width:  1.20
    ```
    
    (Extra challenge: what function did I use to make sure that the numbers
    lined up nicely, even though the variable names had different lengths?)

1.  What does this code do? How does it work?

    ```{r, eval = FALSE}
    trans <- list( 
      disp = function(x) x * 0.0163871,
      am = function(x) {
        factor(x, labels = c("auto", "manual"))
      }
    )
    for (var in names(trans)) {
      mtcars[[var]] <- trans[[var]](mtcars[[var]])
    }
    ```
    
## 21.4: For loops vs. functionals

### 21.4.1

1.  Read the documentation for `apply()`. In the 2d case, what two for loops
    does it generalise?

1.  Adapt `col_summary()` so that it only applies to numeric columns
    You might want to start with an `is_numeric()` function that returns
    a logical vector that has a TRUE corresponding to each numeric column.
    
## 21.5: The map functions

### 21.5.3

1.  Write code that uses one of the map functions to:

    1. Compute the mean of every column in `mtcars`.
    1. Determine the type of each column in `nycflights13::flights`.
    1. Compute the number of unique values in each column of `iris`.
    1. Generate 10 random normals for each of $\mu = -10$, $0$, $10$, and $100$.

1.  How can you create a single vector that for each column in a data frame
    indicates whether or not it's a factor?

1.  What happens when you use the map functions on vectors that aren't lists?
    What does `map(1:5, runif)` do? Why?
    
1.  What does `map(-2:2, rnorm, n = 5)` do? Why?
    What does `map_dbl(-2:2, rnorm, n = 5)` do? Why?

1.  Rewrite `map(x, function(df) lm(mpg ~ wt, data = df))` to eliminate the 
    anonymous function. 
    
## 21.9 Other patterns of for loops

### 21.9.3

1.  Implement your own version of `every()` using a for loop. Compare it with
    `purrr::every()`. What does purrr's version do that your version doesn't?

1.  Create an enhanced `col_sum()` that applies a summary function to every
    numeric column in a data frame.

1.  A possible base R equivalent of `col_sum()` is:

    ```{r}
    col_sum3 <- function(df, f) {
      is_num <- sapply(df, is.numeric)
      df_num <- df[, is_num]

      sapply(df_num, f)
    }
    ```
    
    But it has a number of bugs as illustrated with the following inputs:
    
    ```{r, eval = FALSE}
    df <- tibble(
      x = 1:3, 
      y = 3:1,
      z = c("a", "b", "c")
    )
    # OK
    col_sum3(df, mean)
    # Has problems: don't always return numeric vector
    col_sum3(df[1:2], mean)
    col_sum3(df[1], mean)
    col_sum3(df[0], mean)
    ```
    
    What causes the bugs?