---
title: "Chapter 25 Many models"
date: "`r paste('Last updated: ', format(Sys.time(), '%Y-%m-%d'))`"
author: "Bryan Shalloway"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

*Make sure the following packages are installed:*  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(nycflights13)
library(babynames)
library(nasaweather)
library(lubridate)
```

# ch. 25: Many modeling

## 25.2: gapminder

### 25.2.5

1.  A linear trend seems to be slightly too simple for the overall trend.
    Can you do better with a quadratic polynomial? How can you interpret
    the coefficients of the quadratic? (Hint you might want to transform
    `year` so that it has mean zero.)

1.  Explore other methods for visualising the distribution of $R^2$ per
    continent. You might want to try the ggbeeswarm package, which provides 
    similar methods for avoiding overlaps as jitter, but uses deterministic
    methods.

1.  To create the last plot (showing the data for the countries with the
    worst model fits), we needed two steps: we created a data frame with
    one row per country and then semi-joined it to the original dataset.
    It's possible avoid this join if we use `unnest()` instead of 
    `unnest(.drop = TRUE)`. How?
    
## 25.4: Creating list-columns

### 25.4.5

1.  List all the functions that you can think of that take a atomic vector and 
    return a list.
    
1.  Brainstorm useful summary functions that, like `quantile()`, return
    multiple values.
    
1.  What's missing in the following data frame? How does `quantile()` return
    that missing piece? Why isn't that helpful here?

    ```{r}
    mtcars %>% 
      group_by(cyl) %>% 
      summarise(q = list(quantile(mpg))) %>% 
      unnest()
    ```

1.  What does this code do? Why might might it be useful?

    ```{r, eval = FALSE}
    mtcars %>% 
      group_by(cyl) %>% 
      summarise_each(funs(list))
    ```

## 25.5: Simplifying list-columns

### 25.5.3

1.  Why might the `lengths()` function be useful for creating atomic
    vector columns from list-columns?
    
1.  List the most common types of vector found in a data frame. What makes
    lists different?