---
title: "Chapter 28 Graphics for communication"
date: "`r paste('Last updated: ', format(Sys.time(), '%Y-%m-%d'))`"
author: "Bryan Shalloway"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

*Make sure the following packages are installed:*  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(nycflights13)
library(babynames)
library(nasaweather)
library(lubridate)
```


# ch. 28: Graphics for communication

## 28.2: Label

### 28.2.1

1.  Create one plot on the fuel economy data with customised `title`,
    `subtitle`, `caption`, `x`, `y`, and `colour` labels.

1.  The `geom_smooth()` is somewhat misleading because the `hwy` for
    large engines is skewed upwards due to the inclusion of lightweight
    sports cars with big engines. Use your modelling tools to fit and display
    a better model.

1.  Take an exploratory graphic that you've created in the last month, and add
    informative titles to make it easier for others to understand.
    
## 28.3: Annotations

### 28.3.1

1.  Use `geom_text()` with infinite positions to place text at the
    four corners of the plot.

1.  Read the documentation for `annotate()`. How can you use it to add a text
    label to a plot without having to create a tibble?

1.  How do labels with `geom_text()` interact with faceting? How can you
    add a label to a single facet? How can you put a different label in
    each facet? (Hint: think about the underlying data.)

1.  What arguments to `geom_label()` control the appearance of the background
    box?

1.  What are the four arguments to `arrow()`? How do they work? Create a series
    of plots that demonstrate the most important options.
    
## 28.4: Scales

### 28.4.4

1.  Why doesn't the following code override the default scale?

    ```{r fig.show = "hide"}
    ggplot(df, aes(x, y)) +
      geom_hex() +
      scale_colour_gradient(low = "white", high = "red") +
      coord_fixed()
    ```

1.  What is the first argument to every scale? How does it compare to `labs()`?

1.  Change the display of the presidential terms by:

    1. Combining the two variants shown above.
    1. Improving the display of the y axis.
    1. Labelling each term with the name of the president.
    1. Adding informative plot labels.
    1. Placing breaks every 4 years (this is trickier than it seems!).

1.  Use `override.aes` to make the legend on the following plot easier to see.

    ```{r, dev = "png", out.width = "50%"}
    ggplot(diamonds, aes(carat, price)) +
      geom_point(aes(colour = cut), alpha = 1/20)
    ```
    

